\bf \texttt{forestecology} package for modeling interspecies competition between trees

### 2021-01-09 etc/paper/forestecology.tex

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  set the context for and purpose of the work
\item
  indicate the approach and the methods
\item
  outline the main results
\item
  identify the conclusions and wider implications
\end{enumerate}

@@

\CommentTok{# Load spatial packages:}
\CommentTok{# devtools::install_github("rvalavi/blockCV")}
\KeywordTok{library}\NormalTok{(blockCV)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(sfheaders)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Both Big Woods & SCBI}
\KeywordTok{bind_rows}\NormalTok{(}
\NormalTok{  bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"Big Woods"}\NormalTok{),}
\NormalTok{  scbi_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_drop_geometry}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }\KeywordTok{select}\NormalTok{(growth) }\OperatorTok{%>%}\StringTok{ }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{Site =} \StringTok{"SCBI"}\NormalTok{)}
\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{ggplot}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ ..density.., }\DataTypeTok{fill =}\NormalTok{ Site)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_histogram}\NormalTok{(}\DataTypeTok{alpha =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{position =} \StringTok{"identity"}\NormalTok{, }\DataTypeTok{binwidth =} \FloatTok{0.05}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Average annual growth in dbh (cm per yr)"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_cartesian}\NormalTok{(}\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\FloatTok{0.5}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ bw_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size,}
  \DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{,}
  \DataTypeTok{verbose =} \OtherTok{FALSE}
\NormalTok{)}

\CommentTok{# Add foldID to each tree}
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}

\CommentTok{# Visualize grid. Why does fold 19 repeat?}
\NormalTok{bw_spatialBlock}\OperatorTok{$}\NormalTok{plots }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{)}

\CommentTok{# Remove empty folds}
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{as.character}\NormalTok{(foldID))}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-1} \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-2} \end{center}

@@

\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }\DataTypeTok{region =}\NormalTok{ scbi_study_region)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{1}\NormalTok{)}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{2}\NormalTok{)}
\NormalTok{scbi_cv_grid <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(fold1, fold2)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df,}
  \DataTypeTok{k =} \DecValTok{2}\NormalTok{,}
  \DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{,}
  \DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
  \CommentTok{# Note new arguments:}
  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{,}
  \DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
\NormalTok{)}

@@

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(focal_vs_comp_bw)}
\CommentTok{## # A tibble: 6 x 10}
\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
\CommentTok{##      <dbl> <fct>    <dbl> <chr>                    <POINT>  <dbl>}
\CommentTok{## 1        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 2        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 3        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 4        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 5        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## 6        1 white_o~  41.2 12                   (8.7 107.5)  0.404}
\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{head}\NormalTok{(focal_vs_comp_scbi)}
\CommentTok{## # A tibble: 6 x 10}
\CommentTok{##   focal_ID focal_sp   dbh foldID                  geometry growth}
\CommentTok{##      <dbl> <fct>    <dbl>  <int>                   <POINT>  <dbl>}
\CommentTok{## 1        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 2        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 3        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 4        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 5        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## 6        4 nysy      13.6      1              (14.2 428.5)  0.103}
\CommentTok{## # ... with 4 more variables: comp_ID <dbl>, dist <dbl>,}
\CommentTok{## #   comp_sp <fct>, comp_basal_area <dbl>}
\end{Highlighting}
\end{Shaded}

@@

Next we fit the following linear model to the dbh of each focal tree.
Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of
\texttt{focal\textquotesingle{}\textquotesingle{}\ species\ group\ \$j\$;\ let\ \$j\ =\ 1,\ \textbackslash{}ldots,\ J\$\ index\ all\ \$J\$\ focal\ species\ groups;\ and\ let\ \$k\ =\ 1,\ \textbackslash{}ldots,\ K\$\ index\ all\ \$K\$}competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{ggplot}\NormalTok{(scbi_growth_df_noCV, }\KeywordTok{aes}\NormalTok{(}\DataTypeTok{x =}\NormalTok{ growth, }\DataTypeTok{y =}\NormalTok{ growth_hat)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{, }\DataTypeTok{color =} \KeywordTok{rgb}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\DecValTok{0}\NormalTok{, }\FloatTok{0.25}\NormalTok{)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{stat_smooth}\NormalTok{(}\DataTypeTok{method =} \StringTok{"lm"}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_abline}\NormalTok{(}\DataTypeTok{slope =} \DecValTok{1}\NormalTok{, }\DataTypeTok{intercept =} \DecValTok{0}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{coord_fixed}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}
    \DataTypeTok{x =} \StringTok{"Observed growth in dbh"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Predicted growth in dbh"}\NormalTok{,}
    \DataTypeTok{title =} \StringTok{"Predicted vs Observed Growth"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

@@

\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_growth_df_noCV }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{st_as_sf}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\CommentTok{# }\AlertTok{TODO}\CommentTok{: Need to investigate missingness}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\KeywordTok{is.na}\NormalTok{(growth_hat)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{error =}\NormalTok{ growth }\OperatorTok{-}\StringTok{ }\NormalTok{growth_hat,}
    \DataTypeTok{error_bin =} \KeywordTok{cut_number}\NormalTok{(error, }\DataTypeTok{n =} \DecValTok{5}\NormalTok{),}
    \DataTypeTok{error_compress =} \KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{<}\StringTok{ }\FloatTok{-0.75}\NormalTok{, }\FloatTok{-0.75}\NormalTok{, }\KeywordTok{ifelse}\NormalTok{(error }\OperatorTok{>}\StringTok{ }\FloatTok{0.75}\NormalTok{, }\FloatTok{0.75}\NormalTok{, error))}
\NormalTok{  ) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ error_compress), }\DataTypeTok{size =} \DecValTok{1}\NormalTok{) }\OperatorTok{+}
\StringTok{  }\KeywordTok{theme_bw}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{scale_color_gradient2}\NormalTok{(}
    \DataTypeTok{low =} \StringTok{"#ef8a62"}\NormalTok{, }\DataTypeTok{mid =} \StringTok{"#f7f7f7"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#67a9cf"}\NormalTok{,}
    \DataTypeTok{name =} \KeywordTok{expression}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\StringTok{"Residual (cm "}\NormalTok{, y}\OperatorTok{^}\NormalTok{\{}\OperatorTok{-}\DecValTok{1}\NormalTok{\}, }\StringTok{")"}\NormalTok{)),}
    \DataTypeTok{breaks =} \KeywordTok{seq}\NormalTok{(}\DataTypeTok{from =} \FloatTok{-0.75}\NormalTok{, }\DataTypeTok{to =} \FloatTok{0.75}\NormalTok{, }\DataTypeTok{by =} \FloatTok{0.25}\NormalTok{),}
    \DataTypeTok{labels =} \KeywordTok{c}\NormalTok{(}\StringTok{"< -0.75"}\NormalTok{, }\StringTok{"-0.5"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0"}\NormalTok{, }\StringTok{"0.25"}\NormalTok{, }\StringTok{"0.5"}\NormalTok{, }\StringTok{"> 0.75"}\NormalTok{)}
\NormalTok{  ) }\OperatorTok{+}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x =} \StringTok{"Meter"}\NormalTok{, }\DataTypeTok{y =} \StringTok{"Meter"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/scbi-model-residuals-2-1}

@@

\caption{Spatial distribution of residuals for model applied to SCBI data part 2.}\label{fig:scbi-model-residuals-2}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }\DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}

@@

Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
are passed to \texttt{plot\_posterior\_parameters()} to generate
visualizations of the posterior parameters. These visualizations are
displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
we only plot a subset of the species families.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots <-}\StringTok{ }\KeywordTok{plot_posterior_parameters}\NormalTok{(}
  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw,}
  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots[[}\StringTok{"beta_0"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=0.5\linewidth]{Figures/bw-posterior-viz-beta0-1}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots[[}\StringTok{"beta_dbh"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  set the context for and purpose of the work: models for interspecific
  competition. Robert and spatial crossvalidation
\item
  indicate the approach and the methods: use tidyverse, sf, and
  tidymodels packages.
\item
  outline the main results: Recreate the figure in PLOSOne paper. Code
  would've been way more complicated in base. We don't have to worry
  about how the component functions work, only that sequence/converyor
  belt is correct and output is correct. scientist is abstracted away
  from ugly programming details.
\item
  identify the conclusions and wider implications: the fact truths hold
  was already covered in PLOSOne paper, this package implements this and
  can serve as a blueprint for others who model
\end{enumerate}

@@

\CommentTok{# Load spatial packages:}
\KeywordTok{library}\NormalTok{(blockCV)}
\KeywordTok{library}\NormalTok{(sf)}
\KeywordTok{library}\NormalTok{(sfheaders)}

@@

@@

\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
                      \DataTypeTok{region =}\NormalTok{ bw_study_region)}

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =}\NormalTok{ buffer), }\DataTypeTok{size =} \FloatTok{0.5}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{76}\NormalTok{)}
\NormalTok{bw_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ bw_growth_df, }\DataTypeTok{theRange =}\NormalTok{ cv_fold_size, }\DataTypeTok{k =} \DecValTok{28}\NormalTok{, }\DataTypeTok{xOffset =} \FloatTok{0.5}\NormalTok{, }
  \DataTypeTok{yOffset =} \DecValTok{0}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\hypertarget{add-foldid-to-each-tree}{%
\section{Add foldID to each tree}\label{add-foldid-to-each-tree}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{foldID)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{# Visualize grid. Why does fold 19 repeat?}
\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{()) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_growth_df }\OperatorTok{%>%}\StringTok{ }\KeywordTok{sample_frac}\NormalTok{(}\FloatTok{0.2}\NormalTok{), }
          \KeywordTok{aes}\NormalTok{(}\DataTypeTok{col =} \KeywordTok{factor}\NormalTok{(foldID)), }\DataTypeTok{size =} \FloatTok{0.1}\NormalTok{, }\DataTypeTok{show.legend =} \OtherTok{FALSE}\NormalTok{) }\OperatorTok{+}\StringTok{ }
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ bw_spatialBlock}\OperatorTok{$}\NormalTok{blocks }\OperatorTok{%>%}\StringTok{ }\KeywordTok{st_as_sf}\NormalTok{(), }
               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ folds))}
\end{Highlighting}
\end{Shaded}

{\centering \includegraphics[width=1\linewidth]{Figures/bw-define-cv-folds-with-trees-1}

\caption{Inspect blocks closely.}\label{fig:bw-define-cv-folds-with-trees}
\end{figure}

\hypertarget{remove-empty-folds}{%
\section{Remove empty folds}\label{remove-empty-folds}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bw_growth_df <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{foldID }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\DecValTok{19}\NormalTok{, }\DecValTok{23}\NormalTok{, }\DecValTok{21}\NormalTok{, }\DecValTok{17}\NormalTok{, }\DecValTok{8}\NormalTok{, }\DecValTok{19}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{foldID =} \KeywordTok{as.character}\NormalTok{(foldID))}
\end{Highlighting}
\end{Shaded}

@@

\NormalTok{scbi_growth_df <-}\StringTok{ }\NormalTok{scbi_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{add_buffer_variable}\NormalTok{(}\DataTypeTok{direction =} \StringTok{"in"}\NormalTok{, }\DataTypeTok{size =}\NormalTok{ max_dist, }
                      \DataTypeTok{region =}\NormalTok{ scbi_study_region)}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fold1 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{0}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{1}\NormalTok{)}
\NormalTok{fold2 <-}\StringTok{ }\KeywordTok{rbind}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{300}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{300}\NormalTok{, }\DecValTok{600}\NormalTok{), }\KeywordTok{c}\NormalTok{(}\DecValTok{150}\NormalTok{, }\DecValTok{600}\NormalTok{)) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{sf_polygon}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{folds =} \DecValTok{2}\NormalTok{)}
\NormalTok{scbi_cv_grid <-}\StringTok{ }\KeywordTok{bind_rows}\NormalTok{(fold1, fold2)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{scbi_spatialBlock <-}\StringTok{ }\KeywordTok{spatialBlock}\NormalTok{(}
  \DataTypeTok{speciesData =}\NormalTok{ scbi_growth_df, }\DataTypeTok{k =} \DecValTok{2}\NormalTok{, }\DataTypeTok{verbose =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{showBlocks =} \OtherTok{FALSE}\NormalTok{,}
  \CommentTok{# Note new arguments:}
  \DataTypeTok{selection =} \StringTok{"systematic"}\NormalTok{, }\DataTypeTok{blocks =}\NormalTok{ scbi_cv_grid}
\NormalTok{)}

@@

\KeywordTok{ggplot}\NormalTok{() }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf_text}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_growth_df, }
               \KeywordTok{aes}\NormalTok{(}\DataTypeTok{label =}\NormalTok{ foldID, }\DataTypeTok{col =}\NormalTok{ buffer)) }\OperatorTok{+}
\StringTok{  }\KeywordTok{geom_sf}\NormalTok{(}\DataTypeTok{data =}\NormalTok{ scbi_cv_grid, }\DataTypeTok{fill =} \StringTok{"transparent"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

Next we fit the following linear model to the dbh of each focal tree.
Let \(i = 1, \ldots, n_j\) index all \(n_j\) trees of ``focal'' species
group \(j\); let \(j = 1, \ldots, J\) index all \(J\) focal species
groups; and let \(k = 1, \ldots, K\) index all \(K\) ``competitor''
species groups. We modeled the growth in diameter per year \(y_{ij}\)
(in centimeters per year) of the \(i^{th}\) tree of focal species group
\(j\) as a linear model \(f\) of the following covariates
\(\vec{x}_{ij}\)

@@

@@

\caption{Predicted versus observed growth.}\label{fig:scbi-model-residuals}
\end{figure}

@@

{\centering \includegraphics[width=0.8\linewidth]{Figures/scbi-model-residuals-2-1}

@@

\caption{Spatial distribution of residuals for model applied to SCBI data.}\label{fig:scbi-model-residuals-2}
\end{figure}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{focal_vs_comp_bw <-}\StringTok{ }\NormalTok{bw_growth_df }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{sp =}\NormalTok{ family) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{create_focal_vs_comp}\NormalTok{(}\DataTypeTok{max_dist =}\NormalTok{ max_dist, }\DataTypeTok{cv_grid_sf =}\NormalTok{ bw_cv_grid, }
                       \DataTypeTok{id =} \StringTok{"treeID"}\NormalTok{)}

@@

Now the posterior parameter outputs of \texttt{fit\_bayesian\_model()}
are passed to \texttt{plot\_bayesian\_model\_parameters()} to generate
visualizations of the posterior parameters. These visualizations are
displayed in Figure 5 of \citet{allen_permutation_2020}. For simplicity
we only plot a subset of the species families.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_bw <-}\StringTok{ }\KeywordTok{plot_bayesian_model_parameters}\NormalTok{(}
  \DataTypeTok{posterior_param =}\NormalTok{ posterior_param_bw,}
  \DataTypeTok{sp_to_plot =} \KeywordTok{c}\NormalTok{(}\StringTok{"cornaceae"}\NormalTok{, }\StringTok{"fagaceae"}\NormalTok{, }\StringTok{"hamamelidaceae"}\NormalTok{, }\StringTok{"juglandaceae"}\NormalTok{, }
                 \StringTok{"lauraceae"}\NormalTok{, }\StringTok{"rosaceae"}\NormalTok{, }\StringTok{"sapindaceae"}\NormalTok{, }\StringTok{"ulmaceae"}\NormalTok{)}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_bw[[}\StringTok{"beta_0"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}

@@

{\centering \includegraphics[width=1\linewidth]{Figures/bw-posterior-viz-beta0-1}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{posterior_plots_bw[[}\StringTok{"beta_dbh"}\NormalTok{]]}
\end{Highlighting}
\end{Shaded}
